---
title: "HW 2"
author: "Ethan TUrner"
date: "2/9/2023"
output: 
  html_document:
    number_sections: true
---

This homework is meant to illustrate the methods of classification algorithms as well as their potential pitfalls.  In class, we demonstrated K-Nearest-Neighbors using the `iris` dataset.  Today I will give you a different subset of this same data, and you will train a KNN classifier.  

```{r, echo = FALSE}
set.seed(123)
library(class)

df <- data(iris) 

normal <-function(x) {
  (x -min(x))/(max(x)-min(x))   
}

iris_norm <- as.data.frame(lapply(iris[,c(1,2,3,4)], normal))

subset <- c(1:45, 58, 60:70, 82, 94, 110:150)
iris_train <- iris_norm[subset,] 
iris_test <- iris_norm[-subset,] 

iris_target_category <- iris[subset,5]
iris_test_category <- iris[-subset,5]


```

#
Above, I have given you a training-testing partition.  Train the KNN with $K = 5$ on the training data and use this to classify the 50 test observations.  Once you have classified the test observations, create a contingency table -- like we did in class -- to evaluate which observations your algorithm is misclassifying.   

```{r}
set.seed(123)


pr<- knn(iris_train, iris_test, cl = iris_target_category, k = 5)

tab <- table(pr, iris_test_category)


tab

accuracy<- function(x){
  sum(diag(x))/sum(rowSums(x))*100
}

accuracy(tab)



```

#

Discuss your results.  If you have done this correctly, you should have a classification error rate that is roughly 20% higher than what we observed in class.  Why is this the case? In particular run a summary of the `iris_test_category` as well as `iris_target_category` and discuss how this plays a role in your answer.  



I do indeed have an error rate of 22%, which is about 20% higher than seen in class. As seen in the summaries below, versicolor is overrepresented in our test category when contrasted to the target category. Setosa and virginica are underrepresented, meaning the sheer amount of versicolor adjacent to virginica data points causes many virginica to possess more versicolor neighbors. 

```{r}
summary(iris_test_category)

summary(iris_target_category)
```


#

Build a github repository to store your homework assignments.  Share the link in this file.  

*STUDENT INPUT*
